An Effective Model of HL-60 Differentiation

### 2017-01-27 Paper_v2.tex
E2F cell cycle

\subsubsection*{Cell cycle arrest.}
A defining feature of ATRA-induced differentiation is cell cycle arrest.
We did not include a description of the restriction point logic, however, we do have a model of E2F transcription.
We developed a heuristic

>>>>

\subsubsection*{Effective cell cycle arrest model.}
A defining feature of ATRA-induced differentiation is G1/G0 cell cycle arrest followed by differentiation.
While we included E2F transcription, we do not consider a detailed model of restriction point regulation.
Instead, we formulated an effective model of the fraction of cells undergoing ATRA-induced cell cycle arrest, $\mathcal{A}$ as:
\begin{equation}\label{eq:a-eqn}
	\frac{d\mathcal{A}}{dt} = \epsilon_{1}\left(\mathrm{E2F}\right)^{\epsilon_{2}}
\end{equation}
The parameters $\epsilon_{1}$ and $\epsilon_{2}$ were estimated from experimental cell distribution measurements by \texttt{log} transforming Eq. \eqref{eq:a-eqn}:
\begin{eqnarray}
\log\epsilon_{1}+\epsilon_{2}\log\left(\mathrm{E2F}_{1}\right) & = & \log{\dot{\mathcal{A}}_{21}} \\
\log\epsilon_{1}+\epsilon_{2}\log\left(\mathrm{E2F}_{2}\right) & = & \log{\dot{\mathcal{A}}_{32}}
\end{eqnarray}
where subscripts denote the experimental time point (1 = 24 hr, 2 = 48 hr, and 3 = 72 hr), $\mathrm{E2F}_{i}$ denotes the $\mathrm{E2F}$ abundance at time point $i$,
and $\dot{\mathcal{A}}_{ij}$ denotes the time derivative of the fraction of arrested cells (estimated from data).
We calculated candidate values for $\epsilon_{j}$ by solving the log-transformed system of equations for 10 biological replicates.
